---
title: "Assignment2_Writeup"
format: html
editor: visual
---

```{r}
#| echo: false
#read in data and load necessary packages 
library(gtsummary)
library(dplyr)
library(boot)
library(ggplot2)

data <- read.csv("raw-data/cohort.csv")
#summary(data)

```

## Descriptive statistics

This simulated data includes 5,000 observations with the following variables:

-   smoke: binary variable; 1 - smoking; 0 - no smoking

-   Female: binary variable: 1 - female; 0 - male

-   Age: Integer ranging from 18 - 72

-   Cardiac: binary variable; 1 - cardiac event; 0 - no cardiac event

-   Cost: in UDS, integer ranging from \$8,012 to \$11,370

```{r}
tbl_summary(data)
```

## Prediction Algorithm

Goal: Our goal is to predict the cost given information about smoking history, sex, age, and cardiac. We will train the model on 70% of the data, and withhold 30% for testing.

```{r}
set.seed(435)
training <- sample(seq_len(nrow(data)), size = 3500)
df_train <- data[training, ]
df_test <- data[-training, ]

m1 <- glm(cost ~ smoke + female + age + cardiac, 
       data = df_train)
summary(m1)

df_test$predicted <- predict.glm(m1, newdata = df_test)
df_test$difference <- df_test$cost - df_test$predicted


```

## Figures

We plot two figures:

1.  We plot the expected value on the x-axis versus the predicted value along the y-axis.
2.  In the second figure, we also plot the difference between the true values and the predicted values.

```{r}
ggplot(df_test, aes(x = cost, y = predicted)) +
      geom_point(size = 2) +
      theme_minimal()


hist(df_test$difference,
     xlab = "Difference in actual cost versus predicted ($)",
     ylab = "Count")
```

## Conclusions:

In conclusion, we used the glm package to train a generalized linear model to predict cost from four variables, age, sex, smoking history, and cardiac event. We found that all four of the variables included had a significant coefficient, so we included all four in the model.

Then, we tested the model on data it had not previously seen. In Figure 1, we plotted the true costs versus the predicted costs, and we saw a strong linear relationship between the two. In Figure 2, we plotted a histogram of the differences between the predicted costs and the true costs. We found that those difference were normally distributed around zero.

Taken together, this suggests our model does a reasonably sufficient job at predicting costs using age, sex, smoking history, and cardiac event.

## AI Statement: 

No Generative AI was used for any portion of this assignment.
